State-space PTA

### 2023-06-13 docs/manuscript/paper.tex
paper

@@

\subsection{Discretised model} \label{sec:statespace}
We can express the intrinsic frequency evolution, Eq. \ref{eq:frequency_evolution_expanded}, in a alternative form as,
\begin{equation}
	df = \mathcal{A} f dt + N(t) dt + \sigma dB(t)
	\label{eq:state1}
\end{equation}
where $\mathcal{A} = -\gamma$, $N(t) = \gamma(f_{\rm EM}(0) + \dot{f}_{\rm EM}(0) t) +\dot{f}_{\rm EM}(0)$ and $dB(t)$ denotes increments of Brownian motion (Wiener process). This equation is easily identified as an Ornstein-Uhlenbeck process which has a general solution given by \citep{gardiner2009stochastic},
\begin{equation}
	f(t) = e^{\mathcal{A}t}f(0) + \int_0^t e^{\mathcal{A}(t-t')} N(t') dt' + \int_0^t e^{\mathcal{A}(t-t')} \sigma dB(t')
\end{equation} 
If we move from a solution in continuous time, $t$, to discrete time, $\bar{t} = (t_1, t_2, ...,t_K)$, then
\begin{equation}
	f(t_{i+1}) = F f(t_i) + T_i + \eta_i
\end{equation}
where
\begin{equation}
	F_i = e^{\mathcal{A} (t_{i+1} - t_i)}
\end{equation}
\begin{equation}
	T_i = \int_{t_i}^{t_{i+1}}  e^{\mathcal{A} (t_{i+1} - t')} N(t') dt'
\end{equation}
\begin{equation}
	\eta_i = \int_{t_i}^{t_{i+1}}  e^{\mathcal{A} (t_{i+1} - t')} \sigma dt'
\end{equation}
The discrete solution $f(\bar{t})$ to the intrinsic frequency can be related to the discrete measured frequency via Eq. \ref{eq:main_eq} as,
\begin{equation}
	f_M (\bar{t})= f(\bar{t}) g(\bar{\theta},\bar{t}) + N_M
\end{equation}
where $g(\theta,t)$ can be expressed in a trigonometric form as 
\begin{equation}
		X = 1 - \frac{1}{2} \frac{ H_{ij}q^i q^j }{(1 + \bar{n}\cdot \bar{q}) } \left[ \cos(-\Omega \tau +\Phi_0) - \cos(-\Omega \tau +\Phi_0 + \Omega (1 + \bar{n}\cdot \bar{q})  d) \right]
\end{equation} 
whilst $N_M$ is a Gaussian measurement noise that satisfies 
\begin{equation}
	\langle N_M(t) N_M(t') \rangle = \Sigma^2 \delta(t - t')
\end{equation}
for variance $\Sigma^2$.

>>>>

THe standard method of detecting single source psr is as follows

drawback is does not account for noises, ooptimal estimators etc etc.

@@

\subsection{Discretised model} \label{sec:statespace}
We can express the intrinsic frequency evolution, Eq. \ref{eq:frequency_evolution_expanded}, in a alternative form as,
\begin{equation}
	df = \mathcal{A} f dt + N(t) dt + \sigma dB(t)
	\label{eq:state1}
\end{equation}
where $\mathcal{A} = -\gamma$, $N(t) = \gamma(f_{\rm EM}(0) + \dot{f}_{\rm EM}(0) t) +\dot{f}_{\rm EM}(0)$ and $dB(t)$ denotes increments of Brownian motion (Wiener process). This equation is easily identified as an Ornstein-Uhlenbeck process which has a general solution given by \citep{gardiner2009stochastic},
\begin{equation}
	f(t) = e^{\mathcal{A}t}f(0) + \int_0^t e^{\mathcal{A}(t-t')} N(t') dt' + \int_0^t e^{\mathcal{A}(t-t')} \sigma dB(t')
\end{equation} 
If we move from a solution in continuous time, $t$, to discrete time, $\bar{t} = (t_1, t_2, ...,t_K)$, then
\begin{equation}
	f(t_{i+1}) = F f(t_i) + T_i + \eta_i
\end{equation}
where
\begin{equation}
	F_i = e^{\mathcal{A} (t_{i+1} - t_i)}
\end{equation}
\begin{equation}
	T_i = \int_{t_i}^{t_{i+1}}  e^{\mathcal{A} (t_{i+1} - t')} N(t') dt'
\end{equation}
\begin{equation}
	\eta_i = \int_{t_i}^{t_{i+1}}  e^{\mathcal{A} (t_{i+1} - t')} \sigma dt'
\end{equation}

if we specialise to the case of constatn time sampling

The discrete solution $f(\bar{t})$ to the intrinsic frequency can be related to the discrete measured frequency via Eq. \ref{eq:main_eq} as,
\begin{equation}
	f_M (\bar{t})= f(\bar{t}) g(\bar{\theta},\bar{t}) + N_M
\end{equation}
where $g(\theta,t)$ can be expressed in a trigonometric form as 
\begin{equation}
		X = 1 - \frac{1}{2} \frac{ H_{ij}q^i q^j }{(1 + \bar{n}\cdot \bar{q}) } \left[ \cos(-\Omega \tau +\Phi_0) - \cos(-\Omega \tau +\Phi_0 + \Omega (1 + \bar{n}\cdot \bar{q})  d) \right]
\end{equation} 
whilst $N_M$ is a Gaussian measurement noise that satisfies 
\begin{equation}
	\langle N_M(t) N_M(t') \rangle = \Sigma^2 \delta(t - t')
\end{equation}
for variance $\Sigma^2$.
